---
title: "Project One"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## MT5763 Software and Data Analysis **Due: 8th October**

Please find my GitHub repo here <https://github.com/selsiekiely/MT5763_1_190018562>.

### **Contents**
### Data Wrangling
###Data Visualisation
###Statistical Modelling

##**Data Wrangling**
## We will now data wrangle the BikeSeoul.csv

```{r, include = FALSE}
library(tidyverse)
library(readr)
BikeSeoul <- read_csv("~/Year 3/MT5763/BikeSeoul.csv")
View(BikeSeoul)
BikeWashington <- read_csv("~/Year 3/MT5763/BikeWashingtonDC.csv")
View(BikeWashington)
```

```{r BikeSeoul}
BikeSeoul <- BikeSeoul %>%
  select(- `Visibility (10m)`, -`Dew point temperature(C)`, -`Solar Radiation (MJ/m2)`, 
         -`Rainfall(mm)`, -`Snowfall (cm)`)
BikeSeoul <- BikeSeoul %>%  
  filter(!is.na(`Rented Bike Count`))
BikeSeoul <- BikeSeoul %>% 
  select(-`Functioning Day`)
library(lubridate)
BikeSeoul <- BikeSeoul %>%
  rename(Count = `Rented Bike Count`) %>%
  rename(Temperature = `Temperature(C)`) %>%
  rename(Humidity = `Humidity(%)`) %>%
  rename(WindSpeed = `Wind speed (m/s)`) %>%
  rename(Season = `Seasons`) %>%
  mutate(Date = as.Date(Date, format = "%d/%m/%Y"))
BikeSeoul <- BikeSeoul[!(BikeSeoul$Count == 0),]
year(BikeSeoul$Date) <- year(BikeSeoul$Date)
BikeSeoul <- BikeSeoul %>%
  mutate(BikeSeoul, FullDate = make_datetime(year = year(Date), month = month(Date), day = mday(Date), Hour)) %>%
  mutate(Temperature = as.numeric(Temperature))
BikeSeoul$Holiday <- ifelse(BikeSeoul$Holiday == "No Holiday", "No", "Yes")
yes_no_levels <- c("Yes", "No")
season_levels <- c("Spring", "Summer", "Autumn", "Winter")
BikeSeoul <- BikeSeoul %>%
  mutate(Holiday = factor(Holiday, levels = yes_no_levels),
         Season = factor(Season, levels = season_levels)) %>%
   arrange(Season, Holiday)
View(BikeSeoul)
```

```{r BikeWashington}
BikeWashington <- BikeWashington %>%
  select(-instant, -yr, -mnth, -weekday, -workingday, -weathersit, -atemp, -casual, -registered)
BikeWashington <- BikeWashington %>%
  rename(Date = dteday) %>%
  rename(Count = cnt) %>%
  rename(Hour = hr) %>%
  rename(Temperature = temp) %>%
  rename(Humidity = hum) %>%
  rename(WindSpeed = windspeed) %>%
  rename(Season = season) %>%
  rename(Holiday = holiday)
BikeWashington$Humidity <- paste(100*BikeWashington$Humidity)
BikeWashington$Temperature <- paste(47*BikeWashington$Temperature - 8)
BikeWashington$WindSpeed <- paste(67*(18/5)*BikeWashington$WindSpeed)
BikeWashington$Holiday <- ifelse(BikeWashington$Holiday == 0, "No", "Yes")
BikeWashington$Season <- ifelse(BikeWashington$Season == 1, "Winter", ifelse(BikeWashington$Season == 2, "Spring", ifelse(BikeWashington$Season == 3, "Summer", "Autumn")))
season_levels <- c("Spring", "Summer", "Autumn", "Winter")
yes_no_levels <- c("Yes", "No")
BikeWashington <- BikeWashington %>%
  mutate(Holiday = factor(Holiday, levels = yes_no_levels),
         Season = factor(Season, levels = season_levels)) %>%
   arrange(Season, Holiday)
BikeWashington$Date <- as.Date(BikeWashington$Date)
BikeWashington <- BikeWashington %>%
  mutate(BikeWashington, FullDate = make_datetime(year = year(Date), month = month(Date), day = mday(Date), Hour)) %>% 
  mutate(Temperature = as.numeric(Temperature)) %>%
  mutate(Humidity = as.numeric(Humidity)) %>%
  mutate(WindSpeed = as.numeric(WindSpeed))
View(BikeWashington)
```
# **Data Visualisation**
#### Before we start lets create a conjoined data set that contains a column of location
```{r, include = FALSE}
BikeSeoul <- BikeSeoul %>%
  mutate(BikeSeoul, Location = "Seoul")
BikeWashington <- BikeWashington %>%
  mutate(BikeWashington, Location = "Washington DC")
BikeWashingtonAndSeoul <- full_join(BikeSeoul, BikeWashington)
View(BikeWashingtonAndSeoul)
```
## How does air temperature vary over the course of a year?
### Create Date function to map date to a real number using lubridate package to allow plotting.

```{r, include= FALSE}
library(dplyr)
library(ggplot2)
date_map <- function(Date){
  hr <- as.character(hour(Date))
  mt <- as.character(month(Date))
  dy <- as.character(day(Date))
  if(length(dy) == 1){
    dy <- paste("0", dy, sep = "")
  }
  if(length(mt) == 1){
    mt <- paste("0", mt, sep = "")
  }
  if(length(hr) == 1){
    hr <- paste("0", hr, sep = "")
  }
  return(as.numeric(paste(mt,dy,hr, sep = "")))
}
FullDate_columnseoul <- date_map(BikeSeoul$FullDate)
#BikeSeoul %>% count(year(Date), Temperature) %>% summary()
#BikeWashington %>% count(year(Date), Temperature) %>% summary()
ggplot() + 
  geom_point(BikeSeoul, mapping = aes(x = FullDate_columnseoul, y = Temperature), color='green')# + 
  #geom_point(BikeWashington, mapping = aes(x = month(Date), y = Temperature), color='red') +
  xlab("Year Progress in Months") + ylab("Temperature (Degrees Celsius)") +
    ggtitle("Temperature change over the course of the year")
#ggplot(First_graph_data) +
 # geom_point(aes(x = year(Date), y = Temperature))
```
## How do Seasons affect the average number of rented bikes?
### I will find a daily average for rented bikes and compile this data to compare against Seasons in a plot.
```{r, include= FALSE}
Seoul_daily_count <- BikeSeoul %>%
  group_by(Date) %>%
  summarise(mn = mean(Count))
Washington_daily_count <- BikeWashington %>%
  group_by(Date) %>%
  summarise(mn = mean(Count))
head(Washington_daily_count)

ggplot() +
  geom_point(Seoul_daily_count, mapping = aes(x = month(Date), y = mn), color='blue') +
  geom_point(Washington_daily_count, mapping = aes(x = month(Date), y = mn), color='red') +
  xlab("Date") + ylab("Average Bike Rent") +
    ggtitle("Average Bike Rent over Season change")
```
## Do Holidays increase or decrease the demand for rented bikes?
### Conjoin data sets and compare Yes/No holiday to month and rent count.

```{r, include=FALSE}
ggplot(BikeSeoul) +
  geom_point(mapping = aes(x = Date, y = Count))+
  facet_wrap(~Holiday)
ggplot(BikeWashington) +
  geom_point(mapping = aes(x = Date, y = Count)) +
  facet_wrap(~Holiday)
  
```
## How does the time of day affect the demand for rented bikes?
### I will use lubridate to pick out the time of day and plot this against count.

```{r, include=FALSE}
ggplot() +
  geom_point(BikeSeoul, mapping = aes(x= hour(FullDate), y = Count), color = 'green') +
  geom_point(BikeWashington, mapping = aes(x = hour(FullDate), y = Count), color = 'red')
```
## Is there an association between Bike demand and the three meteorological variables?
### Facet wrap Count vs. Three Meteorological Variables for each BikeSeoul and BikeWashington

```{r, include= FALSE}
ggplot(BikeSeoul) +
  geom_point(mapping = aes(x = Temperature, y = Count, color = 'blue'))+
  geom_point(mapping = aes(x = WindSpeed, y = Count, color = 'green'))+
  geom_point(mapping = aes(x = Humidity, y = Count, color = 'red'))
  
ggplot(BikeWashington) +
  geom_point(mapping = aes(x = Date, y = Count)) +
  facet_wrap(~Holiday)

```
## **Statistical Modelling**
### Firstly, we are required to fit a linear model

```{r, include = FALSE}
first_lm_Seoul <- lm(log(Count) ~ Season + Temperature + Humidity + WindSpeed, data = BikeSeoul)
first_lm_Washington <- lm(log(Count) ~ Season + Temperature + Humidity + WindSpeed, data = BikeWashington)
summary(first_lm_Seoul)
summary(first_lm_Washington)
```
## Next we will calculate Confidence Intervals for the regression coefficients
```{r, include=FALSE}
confint(first_lm_Seoul, level = 0.97)
confint(first_lm_Washington, level = 0.97)
```
## Calculate 90% **Prediction** intervals for Air Temperature(0 degrees celsius), Light Wind (0.5 m/s), Humidity 20%

```{r, include=FALSE}
aim_frame <- data.frame(Season = "Winter",
                        Temperature = 0,
                        WindSpeed = 0.5,
                        Humidity = 20)
predict(first_lm_Seoul, newdata = aim_frame, interval = "prediction")
predict(first_lm_Washington, newdata = aim_frame, interval = "prediction")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

